Summer 2025 ToDo

### 2025-12-03 Relations/Paper/pldiSubmission/Implications.tex
Created a submission folder to make addressing potential reviewer comments easier.

This section examines the hierarchy of ARS properties.
In programming language theory, $\SN$ and $\CR$ constitute the essential properties that
jointly establish completeness.
Consequently, our analysis centers on two primary categories of properties: those associated
with termination and those associated with confluence.

@@

\begin{definition}\label{def:mf} Let $a \in A$. \hfill
    \begin{description}
        \item[\bq \ule{$a \in \MF_R$}] \emph{$a$ is a minimal form} if $a \mstep b \implies b \mstep a$ for any $b \in A$.
        \item[\bq \ule{$a \in \WM_R$}] \emph{$a$ is weakly minimalizing} if $a \mstep b$ for some $b \in \MF_R$.
        \item[\bq \ule{$a \in \SM_R$}]  \emph{$a$ is strongly minimalizing} if either $a \in MF_R$, or every element one $R$ step from $a$ is strongly minimalizing. (This
        definition is to be understood inductively).
        \item[\bq \ule{$a \in \SMseq_R$}] \emph{$a$ is sequentially strongly minimalizing} if for every $R$-increasing sequence $s$ starting at $a$, there exists
        an $i \in \nat$ such that $s (i) \in \MF$.
        \item[\bq \ule{$a \in \MP_R$}] \emph{$a$ has the minimal form property} if $c \bmstep a \mstep b \implies c \mstep b$ for any $b \in \MF$.
    \end{description}
\end{definition}

@@

The implications in this diagram are formalized in \texttt{Implications.agda}.
The confluence hierarchy requires no classical assumptions, however this is not the case for the hierarchy of terminating properties.
The implication from $\SN$ to $\WN$ relies on the classical decidability of the property
of being $R$-minimal, as demonstrated in our formalization:

@@

The same property is similarly required in order to show that $\SM$ implies $\WM$.

@@

\end{proposition}
\begin{proof}
    See \texttt{module Normalizing-Implications} in \texttt{Implications.agda},\\
    \verb|MF∧WN↔NF : ∀ {x} → MF x × WN x ↔ NF x|\\
    \verb|MF∧SN↔NF : ∀ {x} → MF x × SN x ↔ NF x| \qedhere
\end{proof}

@@

Having established the hierarchies of ARS properties, we now examine which property combinations are sufficient to reverse these implications and achieve completeness. The following tables illustrate, for each property combination, whether confluence (left box) and strong normalization (right box) are obtained. Where properties cannot be derived, we provide counterexamples demonstrating the necessity of additional conditions.

@@

\begin{remark} The following are the key takeaways from the implication tables. \hfill
    \begin{enumerate}
        \item The property $\gWCR$ can sometimes be a substitute for the classical
        property required for $\SN \implies \WN$ (and similarly $\SM \implies \WM$).
        \item $\gWN \land \gUNto \implies \gCR$ does not require any classical assumptions
         or $\gWCR$ for the implication to $\gCR$, unlike $\gSN \land \gUNto$.
        \item $\gWN \land \gUNto \implies \gCR$ only holds when the properties are global. Counterexample~\ref{CE:4} shows that $\WCR \land \WN \land \UNto \nRightarrow \CR$.
        \item We have a generalization of Newman's Lemma as $\SM \land \gWCR \implies \CR$. This is explored in Subsection \ref{subsec:newnewman}.
        \item Both $\SM$ and $\WN$ are required for $\SN$, this connection is explored in Subsection \ref{subsec:SMWNSN}.
    \end{enumerate}
\end{remark}

@@

% The tables show a number of changes when moving from properties holding locally to globally.
% Globally $\WN \land \UNto \implies \CR$ holds whereas it does not locally. Similarly, $\SM \land \WN \implies \SN$ holds
% globally but not locally.
\subsection{Relation between $\SM$, $\WN$, and $\SN$} \label{subsec:SMWNSN}
One key takeaway from the tables is that we obtain completeness when we have the properties $\SM$, $\WN$, and $\NP$. We also
obtain completeness when we have $\gWN$ and $\gSM$.
To show that $\gSM \land \gWN \implies \gSN$ we built on our proof that $\SM \land \WN \land \NP \implies \SN$. If $\gWN$ holds
then our proof no longer requires the property $\NP$. This progression of proofs can be seen in the functions:

@@

\subsection{A sufficient condition for $\gSN \to \gWN$}
\label{sec:SNtoWN}
It is perhaps surprising that the intuitively obvious implication from strong
normalization to weak normalization is actually not provable.
A counterexample is furnished in the file \texttt{WFCounters.agda}.
% This fact is related to the subtleties arising in constructive formulation of
% well-foundedness, as we discuss at the end of Section \ref{sec:Well-foundedness}.
At the same time, this implication is valid for all relations which are
1) decidable, and 2) finitely branching.
This is formalized in file \apcomment{Need to fill this in.}

Since these conditions are in fact satisfied by all of the standard
term rewrite systems and lambda calculi, we see that for such ARSs the implication
is automatically valid.

>>>>

The developments of this section are formalized in the files 
\texttt{Properties.agda} and
\texttt{Implications.agda} inside the \texttt{ARS} folder.

This section examines the hierarchy of ARS properties.
In programming language theory, $\gSN$ and $\gCR$ constitute the essential properties that
jointly establish completeness.
Consequently, our analysis centers on two primary categories of properties: those associated
with termination and those associated with confluence.

@@

\begin{definition}\label{def:mf} Let $a \in A$. \hfill
    \begin{description}
        \item[\bq \ule{$a \in \MF_R$}] $a$ is a \bemph{minimal form} if $a \mstep b \implies b \mstep a$ for any $b \in A$.
        \item[\bq \ule{$a \in \WM_R$}] $a$ is \bemph{weakly minimalizing} if $a \mstep b$ for some $b \in \MF_R$.
        \item[\bq \ule{$a \in \SM_R$}]  $a$ is \bemph{strongly minimalizing} if either $a \in MF_R$, or every element one $R$-step from $a$ is strongly minimalizing. (This
        definition is to be understood inductively.)
        \item[\bq \ule{$a \in \SMseq_R$}] $a$ is \bemph{sequentially strongly minimalizing} if for every $R$-increasing sequence $s$ starting at $a$, there exists
        an $i \in \nat$ such that $s (i) \in \MF$.
        \item[\bq \ule{$a \in \MP_R$}] $a$ has the \bemph{minimal form property} if $c \bmstep a \mstep b \implies c \mstep b$ for any $b \in \MF$.
    \end{description}
\end{definition}

@@

The implications in this diagram are formalized in \texttt{Implications.agda}.
The confluence hierarchy requires no classical assumptions, however this is not the case for the hierarchy of terminating properties.
The implication from $\SN$ to $\WN$ relies on the classical decidability of the property
of being $R$-minimal, hence its formalization takes the form

@@

The necessity of assuming a classical axiom for this implication is discussed 
in Subsection \ref{sec:SNtoWN} and
Section \ref{sec:Well-foundedness}.
The same property is similarly required in order to show that $\SM$ implies $\WM$.

@@

\end{proposition}
\begin{proof}
    See module \texttt{Normalizing-Implications} in \texttt{Implications.agda},\\
    \verb|MF∧WN↔NF : ∀ {x} → MF x × WN x ↔ NF x|\\
    \verb|MF∧SN↔NF : ∀ {x} → MF x × SN x ↔ NF x| \qedhere
\end{proof}

@@

Having established the hierarchies of ARS properties, we now examine which property combinations are sufficient to reverse these implications and achieve completeness. The following tables illustrate, for each property combination, whether confluence (left box) and strong normalization (right box) are obtained. Where these properties cannot be derived, we provide counterexamples demonstrating the necessity of additional conditions; these are labeled as CE-X.

@@

\begin{remark} The following are the key takeaways from the implication tables. \hfill
    \begin{enumerate}
        \item The property $\gWCR$ can sometimes be a substitute for the classical
        property required for $\SN \implies \WN$ (and similarly $\SM \implies \WM$).
        \item $\gWN \land \gUNto \implies \gCR$ does not require any classical assumptions
         or $\gWCR$ for the implication to $\gCR$, unlike $\gSN \land \gUNto$.
        \item $\gWN \land \gUNto \implies \gCR$ only holds when the properties are global. Counterexample~\ref{CE:4} shows that $\WCR \land \WN \land \UNto \nRightarrow \CR$.
        \item Both $\SM$ and $\WN$ are required for $\SN$, see Subsection \ref{subsec:SMWNSN}.
        \item We have a generalization of Newman's Lemma as $\SM \land \gWCR \implies \CR$.
    \end{enumerate}
\end{remark}

@@

% The tables show a number of changes when moving from properties holding locally to globally.
% Globally $\WN \land \UNto \implies \CR$ holds whereas it does not locally. Similarly, $\SM \land \WN \implies \SN$ holds
% globally but not locally.
\subsection{Relation between $\SM$, $\WN$, and $\SN$} \label{subsec:SMWNSN}
One key takeaway from the tables is that we obtain completeness when we have the properties $\SM$, $\WN$, and $\NP$. We also
obtain completeness when we have $\gWN$ and $\gSM$.
To show that $\gSM \land \gWN \implies \gSN$ we build on our proof that $\SM \land \WN \land \NP \implies \SN$. If $\gWN$ holds
then our proof no longer requires the property $\NP$. This progression of proofs can be seen in the functions:

@@

\subsection{A sufficient condition for $\gSN \to \gWN$}
\label{sec:SNtoWN}
It is perhaps surprising that the intuitively obvious implication from strong
normalization to weak normalization is actually not provable.
Indeed, the file \texttt{ARS/Examples.agda} shows that assuming this implication 
for all relations on the natural numbers derives excluded middle.
% This fact is related to the subtleties arising in constructive formulation of
% well-foundedness, as we discuss at the end of Section \ref{sec:Well-foundedness}.
At the same time, this implication is valid for all relations which are
1) decidable, and 2) finitely branching.
This is formalized in the file \texttt{ARS/Implications.agda} as

\verb|dec∧FB→SN⊆WN : R isDec → R isFB → SN ⊆ WN|

Since these conditions are in fact satisfied by all of the standard
term rewrite systems and lambda calculi, we see that for the ARSs they induce, the implication
is valid.
